Ecosystem size and complexity dictate riverine biodiversity

### 2021-03-15 manuscript_sciad.Rmd
edits: methods, minor reorganization
edits: maintext, minor edits
edits: maintext, minor edit

We found decisive support for the global model in explaining $\gamma$ diversity (Bayes factor = `r round(fit$gamma$bayes_factor, 1)`) despite the substantial difference in fish fauna between the study regions. The estimated $\gamma$ diversity increased with increasing watershed area (ecosystem size) and branching probability (ecosystem complexity), and these patterns were consistent across the study regions (**Table 1 and Figure 4**). In particular, the effect of ecosystem complexity was striking in its magnitude. Average predictive comparisons [@gelmanAveragePredictiveComparisons2007] revealed that an expected increase of $\gamma$ diversity per 0.1 branching probability was `r op(ap_est$sim_estimate[2] * 0.1)` $\pm$ `r op(ap_est$sim_se[2] * 0.1)` species (**Materials and Methods** and **Supplementary Materials**); this level of increase in $\gamma$ diversity requires `r round(ap_est$sim_estimate[2] * 0.1 / ap_est$sim_estimate[1])` km^2^ in the watershed area (see **Table S7** for the estimated average predictive comparisons). This result is reasonable because numerous studies have shown that tributaries are the primary source of environmental heterogeneity in river networks [@brownDendriticNetworkStructure2010; @petersonModellingDendriticEcological2013; @mooreEmergentStabilityLarge2015; @teruiMetapopulationStabilityBranching2018; @isaakSlowClimateVelocities2016; @mcguireNetworkAnalysisReveals2014; @abbottUnexpectedSpatialStability2018] - the sole theoretical assumption required for the dual control of $\gamma$ diversity to emerge. Further, the consistent patterns across geographic regions with distinct fish fauna suggest that ecological traits of species in metacommunities may have a minor effect on the scaling relationships, as predicted by our theory.

@@

@@

We used `r n_scenario` sets of parameter combinations (i.e., simulation scenarios) based on the results of extensive sensitivity analysis (**Supplementary Materials**). The parameter combinations comprises four landscape and eight ecological scenarios. Landscape scenarios distinguish spatial patterns of habitat heterogeneity, while ecological scenarios change ecological traits of simulated species. For landscape scenarios, we used four combinations of environmental variation at headwaters ($\sigma_h = 0.01, 1$) and the degree of local environmental noise ($\sigma_l = 0.01, 1$). The inequality of $\sigma_h$ and $\sigma_l$ controls the relationship between branching complexity and habitat heterogeneity (**Figures 1A and 1B**). For ecological scenarios, we varied three ecological parameters that are relevant for dispersal and interspecific competition: dispersal distance ($\theta =$ `r unique(dat_sim$theta)`), dispersal probability ($p_d =$ `r unique(dat_sim$p_dispersal)`), and the maximum value of competition coefficients ($\alpha_{max} =$ `r unique(dat_sim$max_alpha)`). This setup results in eight ecological scenarios with different levels of dispersal and competition.

@@

We considered watersheds as independent replicates of metacommunities if the river network flows into one of the following: the ocean, a large lake/reservoir ($\ge$ 10 km² in the areal area) or a large river that may represent lentic habitats ($\ge$ 5000 km² in the watershed area). Metacommunities with $\ge$ 10 sampling sites were used in our analysis because a small sample size inflates statistical uncertainties in estimating asymptotic species richness (**Supplementary Materials**).

We estimated $\alpha$, $\beta$, and $\gamma$ diversity at each watershed. $\gamma$ diversity is the total species richness at each watershed. Since it is impossible to sample all species present in each watershed, we used asymptotic species richness as a proxy for $\gamma$ diversity. Asymptotic species richness can be interpreted as the estimated true species richness in a given area when estimated using spatial replicates of local community data within a metacommunity [@chaoRarefactionExtrapolationHill2014a]. We estimated asymptotic species richness using sampling sites located in the same watershed (i.e., spatial replicates within a metacommunity) via the R package 'iNEXT' [@hsiehINEXTPackageRarefaction2016]. We assessed the completeness of the observed species richness based on sample coverage, which is the total proportional abundances of observed species to the whole metacommunity [@chaoRarefactionExtrapolationHill2014a]. We excluded watersheds whose sample coverage was below 0.9 because the estimates of asymptotic species richness may contain non-negligible statistical uncertainty. $\alpha$ diversity was estimated by taking an average of local species richness at each site. $\beta$ diversity was estimated by dividing the estimated $\gamma$ diversity by $\alpha$ diversity [@whittakerVegetationSiskiyouMountains1960].

>>>>

We found decisive support for the global model in explaining $\gamma$ diversity (Bayes factor = `r round(fit$gamma$bayes_factor, 1)`) despite the substantial difference in fish fauna between the study regions. The estimated $\gamma$ diversity linearly increased with increasing watershed area (ecosystem size) and branching probability (ecosystem complexity) at a log-log scale (i.e., power-law), and these patterns were consistent across the study regions (**Table 1 and Figure 4**). Thus, watershed-scale riverine biodiversity obeys the general scaling law along the two orthogonal dimensions of ecosystem structure: size and complexity. In particular, the effect of ecosystem complexity was striking in its magnitude. Average predictive comparisons [@gelmanAveragePredictiveComparisons2007] revealed that an expected increase of $\gamma$ diversity per 0.1 branching probability was `r op(ap_est$sim_estimate[2] * 0.1)` $\pm$ `r op(ap_est$sim_se[2] * 0.1)` species (**Materials and Methods** and **Supplementary Materials**); this level of increase in $\gamma$ diversity requires `r round(ap_est$sim_estimate[2] * 0.1 / ap_est$sim_estimate[1])` km^2^ in the watershed area (see **Table S7** for the estimated average predictive comparisons). This result is reasonable because numerous studies have shown that tributaries are the primary source of environmental heterogeneity in river networks [@brownDendriticNetworkStructure2010; @petersonModellingDendriticEcological2013; @mooreEmergentStabilityLarge2015; @teruiMetapopulationStabilityBranching2018; @isaakSlowClimateVelocities2016; @mcguireNetworkAnalysisReveals2014; @abbottUnexpectedSpatialStability2018] - the sole theoretical assumption required for the dual scaling of $\gamma$ diversity to emerge. Further, the consistent patterns across geographic regions with distinct fish fauna suggest that ecological traits of species in metacommunities may have a minor effect on the scaling relationships, as predicted by our theory.

@@

\newpage

@@

We used `r n_scenario` sets of parameter combinations (i.e., simulation scenarios) based on the results of extensive sensitivity analysis (**Supplementary Materials**). The parameter combinations comprises four landscape and eight ecological scenarios. Landscape scenarios distinguish spatial patterns of habitat heterogeneity, while ecological scenarios change ecological traits of simulated species. For landscape scenarios, we used four combinations of environmental variation at headwaters ($\sigma_h = 0.01, 1$) and the degree of local environmental noise ($\sigma_l = 0.01, 1$). The inequality of $\sigma_h$ and $\sigma_l$ controls the relationship between branching complexity and habitat heterogeneity (**Figure 1A and B**). For ecological scenarios, we varied three ecological parameters that are relevant for dispersal and interspecific competition: dispersal distance ($\theta =$ `r unique(dat_sim$theta)`), dispersal probability ($p_d =$ `r unique(dat_sim$p_dispersal)`), and the maximum value of competition coefficients ($\alpha_{max} =$ `r unique(dat_sim$max_alpha)`). This setup results in eight ecological scenarios with different levels of dispersal and competition.

@@

We considered watersheds as independent replicates of metacommunities if the river network flows into one of the following: the ocean, a large lake/reservoir ($\ge$ 10 km² in the areal area) or a large river that may represent lentic habitats ($\ge$ 5000 km² in the watershed area). We estimated $\alpha$, $\beta$, and $\gamma$ diversity at each watershed. $\gamma$ diversity is the total species richness at each watershed. Since it is impossible to sample all species present in each watershed, we used asymptotic species richness as a proxy for $\gamma$ diversity. Asymptotic species richness can be interpreted as the estimated true species richness in a given area when estimated using spatial replicates of local community data within a metacommunity [@chaoRarefactionExtrapolationHill2014a]. We estimated asymptotic species richness using sampling sites located in the same watershed (i.e., spatial replicates within a metacommunity) via the R package 'iNEXT' [@hsiehINEXTPackageRarefaction2016]. Metacommunities with $\ge$ 10 sampling sites were used in our analysis because a small sample size inflates statistical uncertainties in estimating asymptotic species richness (**Supplementary Materials**). We assessed the completeness of the observed species richness based on sample coverage, which is the total proportional abundances of observed species to the whole metacommunity [@chaoRarefactionExtrapolationHill2014a]. We excluded watersheds whose sample coverage was below 0.9 because the estimates of asymptotic species richness may contain non-negligible statistical uncertainty. $\alpha$ diversity was estimated by taking an average of local species richness at each site. $\beta$ diversity was estimated by dividing the estimated $\gamma$ diversity by $\alpha$ diversity [@whittakerVegetationSiskiyouMountains1960].
